RadLib: a radiative property model library for CFD

### 2021-09-30 submissions/CPC_response_to_reviewers_revision2.md

## Reviewer 3

The authors have responded to this reviewer's comments and have revised the manuscript substantially. There are still a
few issues that need to be resolved before it is accepted for publication.

**The authors thank the reviewer for their time and contributions.**

<font color='green'>
1) The capabilities and limitations of RadLib should be more clearly articulated. For example, different property models
   have different temperature ranges of applicability. While the authors specified what happens if the simulation
   temperature goes beyond the appropriate temperature range, they should comment or provide some recommendations on the
   appropriateness of the choice of the property model. For example, WSGG model has a valid temperature range of
   300-2400 K, and RadLib caps the absorption coefficient values at the respective bound if the simulation temperature
   is beyond this range. Can the authors provide some guidance on what is the error expected from such a strategy of
   allowing simulation to continue at ranges beyond the valid range of the property model? If the user knows that the
   simulation temperature is beyond 2400 K, is it recommended to use WSGG? Also, what is the applicable range of
   pressure in the models implemented? Related to this, on repeated readings, this reviewer felt a table to list the
   range of applicability for each model may be useful for understanding the capabilities of RadLib at-a-glance for an
   user.

   **Valid temperature ranges for the WSGG and RCSLW models were noted in their respective sections, and the applicable
   temperature range for the PM model has been added to its section. Additionally, the RCSLW model section specifies the
   range of pressures under which the RCSLW model can be applied. The WSGG model has pressure ranges given as the pressure-path length product, which are now    noted in the paper. A table summarizing these conditions has been added to the beginning of Section 2 as requested. 
   The models included cannot be used reliably outside the temperature and pressure bounds for which the models were developed. In the WSGG model, only the a weights have temperature dependence. The model gives a continuing trend as the temperature crosses above the 2400 K bound, up to at least 3000 K, suggesting use of the model up to 3000 K. This is true for the applicable range of Mr. However, some of the a weights are slightly negative above 2800 K as the upper Mr bound is approached. A statement is added to the paper.**

### 2021-09-30 manuscript/manuscript.tex
DOL: comment 1: updating correct p ranges, and discussion

\begin{table}
\caption{Summary of radiative property models in RadLib.}
\label{t:models}
\centering
\resizebox{\textwidth}{!}{
    \begin{tabular}{l l l l l}
        \hline
        Model & T range (K) & P range (atm) & Gas species & Refs.   \\
        \hline
        PM      & 300\textendash 2500 & 1                 & \ce{H2O}, \ce{CO2}, \ce{CO}, \ce{CH4}  &
        \cite{TNF} \\
        WSGG    & 300\textendash 2400 & 1                 & \ce{H2O}, \ce{CO2} & \cite{Bordbar_2020,
            Bordbar_2014} \\
        RCSLW   & 300\textendash 3000 & 0.1\textendash 50 & \ce{H2O}, \ce{CO2}, \ce{CO} &
        \cite{Solovjov_2017} \\
        \hline
    \end{tabular}
}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The particular version of the model implemented is from the Turbulent Nonpremixed Flames (TNF) workshop \cite{TNF}. This is a long-running collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. The workshop includes a number of laboratory jet flames that have been extensively researched for flame temperature, velocity, and composition profiles, and these serve as target flames for computational scientists to validate simulations and models. The PM model implemented here is the one advocated (and hence widely used) by the TNF workshop. For each species, the PM model provides the absorption coefficient as a fourth or fifth order polynomial fit in one of two temperature ranges to $\kappa(T)$ values computed from the RADCAL database \cite{Grosshandler_1993}.  Ultimately, the model accommodates temperatures between 300 K and 2500 K. See \cite{TNF} for details and references. The PM model here considers the species \ce{CO2}, \ce{H2O}, \ce{CO}, \ce{CH4}.  The last two species have lower absorption than the first two, but are included because \ce{CH4} and syngas (consisting of \ce{CO} and \ce{H2}) are common fuels in combustion applications and so are present in the TNF flames in high concentrations in the fuel stream. (\ce{CO} is also a common species in rich flame regions and is also included in the RCSLW model.)

@@

Weighted sum of gray gases (WSGG) models generate radiation absorption coefficients for use with the spectrally integrated RTE (Eq.~\ref{e:RTE_wsgg}).
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}. This model computes emissivity for varying gas temperature, mixture composition, and path length using the HITEMP 2010 database \cite{Rothman_2010}. The $k_j$ and $a_j$ are then correlated to these results using polynomial fits summarized below. The model considers mixtures of gaseous \ce{CO2} and \ce{H2O}, the two most significant radiatively participating species in most combustion environments, using four gray gases and one clear gas. Absorption coefficients are calculated by
%
\begin{linenomath}
\begin{equation}
    \kappa_j=\sum_{k=0}^{4}d_{j,k}M_r^k,
\end{equation}
\end{linenomath}
%
where $\kappa_j$ is the absorption coefficient for gray gas $j$, $d_{j,k}$ is a correlated model coefficient, and $M_r$ is the molar ratio $x_{\mathrm{H_2O}}/x_{\mathrm{CO_2}}$. The weight factors are calculated by
%
\begin{linenomath}
\begin{equation}
    a_j=\sum_{k=0}^{4}b_{j,k}T_r^k,
\end{equation}
\end{linenomath}
%
where $a_j$ is the weighting factor for gas $j$, $b_{i,j}$ is computed as shown below,  and $T_r$ is a scaled temperature equal to $T/T_{ref}$ with $T_{ref}$=1200 K. The valid temperature range of the model is 300-2400 K; temperatures outside this range use the value at the respective bound. The value of $b_{i,j}$ is calculated by
%
\begin{linenomath}
\begin{equation}
    b_{j,k}=\sum_{i=0}^{4}C_{j,k,i}M_r^i,
\end{equation}
\end{linenomath}
%
where $C_{j,k,i}$ is a correlated model coefficient. The model coefficients $d_{j,k}$ and $C_{j,k,i}$ are provided by Bordbar et~al. \citep{Bordbar_2014,Bordbar_2020} but updated values are used in RadLib.

>>>>

\begin{table}
    \caption{Summary of radiative property models and ranges of applicability. For the WSGG model the pressure range is given in terms of the pressure-path length product.}
\label{t:models}
\centering
\resizebox{\textwidth}{!}{
    \begin{tabular}{l l l l l}
        \hline
        Model & T range                 & P range               & Gas species                            & Refs.   \\
        \hline
        PM      & 300\textendash 2500 K & 1 atm                 & \ce{H2O}, \ce{CO2}, \ce{CO}, \ce{CH4}  & \cite{TNF} \\
        WSGG    & 300\textendash 2400 K & 0.01-60 atm$\cdot$m   & \ce{H2O}, \ce{CO2}                     & \cite{Bordbar_2020, Bordbar_2014} \\
        RCSLW   & 300\textendash 3000 K & 0.1\textendash 50 atm & \ce{H2O}, \ce{CO2}, \ce{CO}            & \cite{Solovjov_2017} \\
        \hline
    \end{tabular}
}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The particular version of the model implemented is from the Turbulent Nonpremixed Flames (TNF) workshop \cite{TNF}. This is a long-running collaboration between experimental and computational researchers in turbulent combustion \citep{Smith_2003,Barlow_2001}. The workshop includes a number of laboratory jet flames that have been extensively researched for flame temperature, velocity, and composition profiles, and these serve as target flames for computational scientists to validate simulations and models. The PM model implemented here is the one advocated (and hence widely used) by the TNF workshop. For each species, the PM model provides the absorption coefficient as a fourth or fifth order polynomial fit in one of two temperature ranges to $\kappa(T)$ values computed from the RADCAL database \cite{Grosshandler_1993}. The model accommodates temperatures between 300 K and 2500 K. See \cite{TNF} for details and references. The PM model here considers the species \ce{CO2}, \ce{H2O}, \ce{CO}, \ce{CH4}.  The last two species have lower absorption than the first two, but are included because \ce{CH4} and syngas (consisting of \ce{CO} and \ce{H2}) are common fuels in combustion applications and so are present in the TNF flames in high concentrations in the fuel stream. (\ce{CO} is also a common species in rich flame regions and is also included in the RCSLW model.)

@@

Weighted sum of gray gases (WSGG) models generate radiation absorption coefficients for use with the spectrally integrated RTE (Eq.~\ref{e:RTE_wsgg}).
RadLib implements the WSGG model presented by Bordbar et al. \citep{Bordbar_2014,Bordbar_2020}. This model computes emissivity for varying gas temperature, mixture composition, and path length using the HITEMP 2010 database \cite{Rothman_2010}. The $k_j$ and $a_j$ are then correlated to these results using polynomial fits summarized below. The model considers mixtures of gaseous \ce{CO2} and \ce{H2O}, the two most significant radiatively participating species in most combustion environments, using four gray gases and one clear gas. Absorption coefficients are calculated by
%
\begin{linenomath}
\begin{equation}
    \kappa_j=\sum_{k=0}^{4}d_{j,k}M_r^k,
\end{equation}
\end{linenomath}
%
where $\kappa_j$ is the absorption coefficient for gray gas $j$, $d_{j,k}$ is a correlated model coefficient, and $M_r$ is the molar ratio $x_{\mathrm{H_2O}}/x_{\mathrm{CO_2}}$. The weight factors are calculated by
%
\begin{linenomath}
\begin{equation}
    a_j=\sum_{k=0}^{4}b_{j,k}T_r^k,
\end{equation}
\end{linenomath}
%
where $a_j$ is the weighting factor for gas $j$, $b_{i,j}$ is computed as shown below,  and $T_r$ is a scaled temperature equal to $T/T_{ref}$ with $T_{ref}$=1200 K. The valid temperature range of the model is 300-2400 K; temperatures outside this range use the value at the respective bound. The $\kappa_j$ values are not temperature dependent. The $a_j$ values show a continuing trend as the temperature increases past 2400 K up to 3000 K, suggesting possible use of the model at these temperatures. However, some of the $a_j$ become slightly negative above $\sim$2800 K as the upper $M_r$ bound is approached. The pressure range of the model is given in terms of the pressure-path length product with values from 0.01-60 atm$\cdot$m. The value of $b_{i,j}$ is calculated by
%
\begin{linenomath}
\begin{equation}
    b_{j,k}=\sum_{i=0}^{4}C_{j,k,i}M_r^i,
\end{equation}
\end{linenomath}
%
where $C_{j,k,i}$ is a correlated model coefficient. The model coefficients $d_{j,k}$ and $C_{j,k,i}$ are provided by Bordbar et~al. \citep{Bordbar_2014,Bordbar_2020} but updated values are used in RadLib.
